### 2024-04-16 products/manuscript/Manuscript.qmd
Manuscript Updates
Added figures, equations, an abstract, and updated some of the sections.

---
title: "An Analysis of Health & Lifestyle Factors Associated with Sleep Quality"
subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../manuscript/dataanalysis-references.bib
editor: 
  markdown: 
    wrap: 72
---

# Summary/Abstract (1)

For our project, we obtained a dataset from Kaggle.com containing 373
observations of 13 variables from a study on different factors that may
affect sleep quality. Variables of interest from the study included
physical activity level, age, sex, occupation, stress, and several more.
In RStudio (version 4.3), we used a combination of packages to create
models suitable for answering our primary question: which variables are
most important for determining an individual's sleep quality and what
effect do they have? After conducting an exploratory data analysis (EDA)
we hypothesized that stress, sleep duration, and occupation would have
the largest impact on sleep quality. We believed that stress and sleep
quality would be negatively correlated, sleep duration and quality would
be positively correlated, and that jobs with longer or more variable
hours would negatively affect sleep quality.

After creating linear models for each variable's impact on sleep quality
and fitting a Random Forest model with Cross-Validation to test variable
importance, we determined that sleep duration, stress level, and age
were the most important factors. Our hypothesis for sleep duration and
stress was supported; however, we did not correctly predict that age
would be one of the most important predictors for sleep quality. Our
hypotheses regarding the effects of sleep duration, stress, and
occupation were all supported as well. Overall, we determined that an
individual's duration of sleep, stress level, and age are all key
factors that can impact sleep quality.

*Feel free to edit this Emma; I tried to include everything needed*

# Introduction (2)

## Background Information

The data was obtained from kaggle.com at this
[link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
It contains data from a study on individuals of different sexes and age
regarding their sleep quality and different variables that may affect
this. We are unsure of where this data set came from and how it was
collected; the publisher on Kaggle has yet to respond to a comment with
this question posted on March 5th 2024. Regardless, it contains
interesting variables that have been shown in previous studies to have
an effect on sleep quality.

For instance, in Sun et al. (2015), a sample of Chinese individuals
showed a correlation between obesity and worsened sleep quality in men
but not women. We are interested in seeing if this trend holds up for
our dataset; given that most of the observations are of men, we expect
to see a correlation between lower sleep scores and obesity in our data.
Furthermore, it has been shown that age and gender have a notable effect
on sleep quality (Madrid-Valero et al. 2016). Women seem to experience a
deterioration in sleep quality as they age; however, while this trend is
still present in men, it is notably less consistent and can vary
dramatically between individuals. We expect to see similar trends in our
dataset, but as stated previously, our observations contain
significantly more men than women. Finally, we wish to test whether or
not certain occupational industries have different effects on sleep
quality. It has been shown that managerial positions tend to have the
worst sleep quality among civilian sector workers while 24 jobs that
have rotating shifts tend to have the worst sleep quality (Luckhaupt et
al. 2010). Our dataset contains observations of mostly
white-collar/service-based jobs, so we expect to find similar trends
upon analysis.

*NEED TO INSERT*

*EMMA'S SOURCES ARE LINKED HERE:*

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC8345503/ --\> blood
pressure and sleep
https://www.apa.org/news/press/releases/stress/2013/sleep --\> sleep
duration & stress on sleep quality
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10503965/#:\~:text=Regular%20physical%20activity%20can%20lead,managing%20sleep%20disorders%20like%20insomnia.
--\> physical activity and sleep quality

{{< pagebreak >}}

# Materials and Methods (3)

## Data and Processing

The dataset originally contained 373 observations for 13 variables
covering a broad spectrum health and lifestyle variables associated with
sleep quality. The curators of the dataset did not provide information
on how or where this data was collected, so we solely relied on the
codebook for variable definitions. There were no missing values or
erroneous variables in the dataset that we needed to remove. The
"Blood.Pressure" variable was transformed to reflect the systolic and
diastolic blood pressure measurements from each subject. We then chose
to add an additional variable, labeled 'cat_bp' to reflect the
categorical blood pressure status of a subject based on the [*American
Heart Association
guidelines*](https://www.heart.org/en/health-topics/high-blood-pressure).
Additionally we created a variable, 'StepsGroup' to determine the
categorical activity level of subjects in the dataset based on these
guidelines provided by 10,000 steps [*10,000
steps.org*](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)*.*
Finally, we created an additional categorical variable,
'PhysicalActivityGroup' based on 30 minute increments, to categorically
represent the level of physical activity reported by a subject. The
outcome of interest is noted as "Quality.of.Sleep" that we will
reference as Sleep Quality or Quality of Sleep throughout this report.

## Variables included in Analyses

Given that there were only 13 original variables, and 17 variables after
final data transformation; simple linear regression models were
performed to determine baseline associations between variables and Sleep
Quality, the outcome of interest. These regression models were created
to determine which variables had the largest impact on Sleep Quality and
drove the rest of the analyses. Simple linear regression models were
fitted for the following variables: BMI, cat_bp, Stress Level, Gender,
Age, and Occupation.

The Occupation variable was then transformed to group the various
occupations into the following groups: Healthcare, Education,
Engineering, Business/Finance, and Science. Prior to performing any
subsequent analyses, we removed the original versions of the variables
we transformed: systolic, diastolic, daily steps, occupation,
Physical.Activity.Level, Heart.Rate, and AgeGroup.

The objective of our analysis was to determine which of these variables
are most important for determining an individual's sleep quality and
what specific effect they have. After conducting an (EDA) we
hypothesized that stress, sleep duration, and occupation would have the
largest impact on sleep quality. We believed that stress and sleep
quality would be negatively correlated, sleep duration and quality would
be positively correlated, and that jobs with longer or more variable
hours would negatively affect sleep quality.

## Model Development

Random Forest Models and Cross-Validation were considered to be the
"best" fit for our outcome of interest, Quality of Sleep. The data were
not split into train/test subsets as the data contains less than 400
observations and many of the values were unique in comparison to the
rest of the data. Prior to constructing this model, a collinearity plot
was constructed to determine any presence of collinearity in our data.
![Collinearity between all variables.](images/collinearity.png)

Stress and Sleep Quality had an absolute value of 0.9 on the correlation
scale. Sleep Duration and Stress were also seen to be strongly
correlated. Stress was still included in our models due to its
well-studied impact on sleep; however, we found it important to note the
strong linear relationship between Stress and Sleep Quality.

A variable importance plot (VIP) was constructed to determine which
variables had the strongest impact in our Random Forest Model. ![The
ranking of each variable\'s importance in determining sleep quality
according to our Random Forest
model.](images/variableimportance.png)Sleep Duration, Stress Level, and
Age had the highest levels of importance in our model. Physical Activity
Level was excluded entirely, which was to be expected given our
exploratory data analysis and simple linear regression models.

## Defining the Models

*DEFINE DECISION TREES. DEFINE BAGGING AND BOOTSTRAPPING. DEFINE RANDOM
FOREST MODELS* *DEFINE CROSS-VALIDATION* *DEFINE IMPORTANCE OF NULL
MODEL TO COMPARE RESULTS*

*INCLUDE CITATIONS OF CHAPTER 9,10,11 OF HMLR LINKED IN "MANY TREE
MODELS" SECTION OF CONTENT FOLDER 11*

## Evaluation of Models

Root Mean Squared Error (RMSE) is a common metric used to evaluate
regression models. The RMSE is formally defined as the square root of
the mean square of all error and is defined by the following formula:
$$\sum_{i=1}^{D}(x_i-y_i)^2$$

It is important to consider that while RMSE is scale dependent, common
practice notes that low RMSE values indicate stronger model performance.

R-squared (R²) is a statistical measure used in regression analyses. It
quantifies the proportion of the variance in a dependent variable that
can be explained by an independent variable within a regression model
and is calculated with the following formula:

$$
[ R^2 = 1 - \frac{{\text{{Unexplained Variation}}}}{{\text{{Total Variation}}}} ]
$$

We used both RMSE and R² to evaluate the performance of our model
compared to a "null model". Null models generate predictions for an
outcome of interest without using any predictors. Instead, they simply
average the values of the outcome of interest to make "predictions".

![The results generated from our RF model. RMSE increases when more
predictors are added to the model. The R-squared value also decreases as
more predictors are added to the model.](images/RFtuningresults.png)

Figure (*name/order figures after we finalize draft*) shows that as we
include more predictors in our model, RMSE increases slightly while the
R-squared value decreases. This indicates that some variables in our
dataset are more important for predicting sleep quality than the others.
Based off of our variable importance graph, we can assume that stress
level, sleep duration, and age are the factors that carry the most
weight when predicting sleep quality.

We went on to evaluate how well our RF model predicted values by
creating an observed vs. predicted plot.

![A comparison of the actual values for sleep quality in our dataset
(observed) compared to the values predicted by our fitted RF
model.](images/observed_vs_predicted.png)

While the predicted values don't align perfectly with the red "ideal
value" line, they fall very close overall, indicating that our model can
sufficiently predict sleep quality using unseen data.

## Software Used for Analyses

This analysis was conducted under R version 4.3.1 on a MacOS operating
system *cite R*. The following R packages were used in the development
of these analyses: *R PACKAGES USED: here, skimr, broom, tidyverse,
ggplot2, dplyr, corrplot, ranger, vip* All processing and analysis code
can be found in the Supplementary Material file.

# Results (4)

## Outcomes of Interest

The 'Quality of Sleep' variable has been selected as our outcome of
interest. It is important to note that this variable is a subjective
measure of a participant's self-reported sleep quality. The following
plot shows that most subjects reported a sleep quality score of 8 (out
of 10).

![Distribution of Sleep Quality Scores in the
dataset.](images/unnamed-chunk-17-1.png)

## Exploratory Findings

## Machine Learning Models

*To Emma: did you want this text in both places you placed it?* A
variable importance plot (VIP) was constructed to determine which
variables had the strongest impact in our Random Forest Model. *INCLUDE
PLOT HERE* Sleep Duration, Stress Level, and Age had the highest levels
of importance in our model. Physical Activity Level was excluded
entirely, which was to be expected given our exploratory data analysis
and simple linear regression models.

*Include comparison of RF Model and null model*

# Discussion (5)

*PART 5*

## Summary

*PART 5*

## Strengths and Limitations

*PART 5*

*Include that all parameters were self-reported and thus subjected to
bias*

## Conclusion

*PART 5*

{{< pagebreak >}}

# References

*INCLUDE INCREMENTALLY -- FINALIZE*

https://link.springer.com/article/10.1007/s11325-015-1193-z --\> obesity
and sleep quality (affects men but not women in China)

https://www.scielosp.org/article/gs/2017.v31n1/18-22/en/?uid=680b54f39b#
--\> effect of age and gender on sleep quality (women worse than men,
sleep quality deteriorates with age in women but is a bit all over the
place for men)

https://academic.oup.com/sleep/article/33/2/149/2454438#77919376 --\>
sleep quality by occupation/industry (managerial positions had the worst
score)