### 2024-04-29 products/manuscript/Manuscript.qmd

---
title: "An Analysis of Health & Lifestyle Factors Associated with Sleep Quality"
subtitle: ""
author: Kevin Kosewick and Emma Hardin-Parker
date: "4-29-2024"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../manuscript/references_sleepdata.bib
editor: 
  markdown: 
    wrap: 72
---

# Summary/Abstract (1)

For our project, we obtained a dataset from
Kaggle.com containing 373 observations of
13 variables from a study on different factors that may affect sleep
quality. Variables of interest from the study included physical activity
level, age, sex, occupation, stress, gender, BMI, blood pressure, and
sleep duration. In [RStudio](https://posit.co/download/rstudio-desktop/)
(version 4.3.1), we used a combination of packages to create models
suitable for answering our primary question: which variables are most
important for determining an individual's sleep quality and what effect
do they have? After conducting an exploratory data analysis (EDA) we
hypothesized that stress, sleep duration, and occupation would have the
largest impact on sleep quality. We believed that stress and sleep
quality would be negatively correlated, sleep duration and quality would
be positively correlated, and that jobs with longer or more variable
hours would negatively affect sleep quality.

After creating linear models for each variable's impact on sleep quality
and fitting a Random Forest model with Cross-Validation to test variable
importance, we determined that sleep duration, stress level, and age
were the most important factors. Our hypothesis for sleep duration and
stress was supported; however, we did not correctly predict that age
would be one of the most important predictors for sleep quality. Our
hypotheses regarding the effects of sleep duration, stress, and
occupation were all supported as well. Overall, we determined that an
individual's duration of sleep, stress level, and age are all key
factors that can impact sleep quality.

# Introduction (2)

## Background Information

The data was obtained from kaggle.com at this
[link](https://www.kaggle.com/datasets/henryshan/sleep-health-and-lifestyle/data).
The data contains observations from a study on individuals of varying
sexes and age regarding their sleep quality and different variables that
may affect this. We are unsure of where this data set came from and how
the data were collected; the publisher on Kaggle has yet to respond to a
comment with this question posted on March 5th 2024. Regardless, it
contains interesting variables that have been shown in previous studies
to have an effect on sleep quality.

For instance, in Sun et al. (2015), a sample of Chinese individuals
showed a correlation between obesity and worsened sleep quality in men
but not women. We are interested in seeing if this trend in present in
our dataset; however, given that most of the observations are of men, we
expect to see a correlation between lower sleep scores and obesity in
our data. Furthermore, it has been shown that age and gender have a
notable effect on sleep quality (Madrid-Valero et al. 2016). Women seem
to experience a deterioration in sleep quality as they age; however,
while this trend is still present in men, it is notably less consistent
and can vary dramatically between individuals. We expect to see similar
trends in our dataset, but as stated previously, our observations
contain significantly more men than women. Finally, we wish to test
whether or not certain occupational industries have different effects on
sleep quality. It has been shown that managerial positions tend to have
the worst sleep quality among civilian sector workers while 24 jobs that
have rotating shifts tend to have the worst sleep quality (Luckhaupt et
al. 2010). Our dataset contains observations of mostly
white-collar/service-based jobs, so we expect to find similar trends
upon analysis.

Additionally, based on the findings of Sun et. al that obesity can
negatively impact sleep quality, we also expect that higher levels of
blood pressure will result in poorer sleep quality. The American Sleep
Foundation states that both short and long sleep duration increase risk
of hypertension in adults (Li & Shang, 2021). This draws interesting
focus to the circular nature of sleep quality and hypertension. Those
with hypertension also face a higher risk of insomnia, a disorder
affecting the ability to fall asleep, maintain sleep, and sleep without
continuous disturbance (Liu et. al, 2022). Short sleep duration is also
positively associated with stress level (APA, 2013). Survey findings
indicate that the average American sleeps 6.7 hours per night, which is
well below the minimum recommended duration seven to nine hours (APA,
2013). Similar to the relationship between hypertension and sleep
quality, as stress level increases, sleep duration decreases thereby
causing poorer sleep quality. Poorer sleep quality then leads to higher
stress, greater risk of hypertension, and shorter sleep duration.

Engaging in physical activity has been suggested to improve both sleep
quality and sleep duration. Regular moderate to intense physical
activity stimulates an increase of melatonin production, a hormone that
regulates sleep-wake cycles (Alnawwar et. al, 2023). Additionally,
moderate to intense exercise has shown to have a positive effect on
those with insomnia (Alnawwar et. al, 2023). Physical activity also
reduces stress and improves both sleep quality and sleep efficiency
(Alnawwar et. al, 2023). However, studies also show that frequent
high-intensity workouts could lead to poorer sleep quality among people
with and without insomnia (Alnawwar et. al, 2023). Studies indicate that
consistent moderate exercise three to four times a week (approximately
150 minutes total) is the optimal weekly regimen to improve sleep
quality and reduce stress. We expect to see similar trends with our data
where increased stress levels negatively impact sleep quality. We expect to see
a positive association between increasing physical activity levels and
sleep quality until the point of "too much" activity.

{{< pagebreak >}}

# Materials and Methods (3)

## Data and Processing

The dataset originally contained 373 observations for 13 variables
covering a broad spectrum of health and lifestyle variables associated
with sleep quality. The curators of the dataset did not provide
information on how or where this data was collected, so we solely relied
on the codebook for variable definitions. We have added this codebook to
the supplementary materials file for ease of access but you can also
find it on the Kaggle page for this dataset. There were no missing
values or erroneous variables in the dataset that we needed to remove.
The "Blood.Pressure" variable was transformed to reflect the individual
systolic and diastolic blood pressure measurements from each subject. We
then chose to add an additional variable, labeled 'cat_bp' to reflect
the categorical blood pressure status of a subject based on the
[American Heart Association
guidelines](https://www.heart.org/en/health-topics/high-blood-pressure).
Additionally we created a variable named 'StepsGroup' to determine the
categorical activity level of subjects in the dataset based on these
guidelines provided by [10,000 steps.org](https://www.10000steps.org.au/articles/healthy-lifestyles/counting-steps/#:~:text=The%20following%20pedometer%20indices%20have%20been%20developed%20to,day%205%20Highly%20active%20is%20more%20than%2012%2C500)
Finally, we created an additional categorical variable called
'PhysicalActivityGroup' to categorically represent the level of physical
activity reported by a subject. The levels are differentiated by 30
minute intervals of weekly physical activity. The outcome of interest is
noted as "Quality.of.Sleep" that we will reference as Sleep Quality or
Quality of Sleep throughout this report.

## Variables included in Analyses

Given that there were only 17 variables after final data transformation,
simple linear regression models were performed to determine baseline
associations between variables and Sleep Quality (the outcome of
interest). These regression models were created to determine which
variables had the largest impact on Sleep Quality and drove the rest of
the analyses. Simple linear regression models were fitted for the
following variables: BMI, cat_bp, Stress Level, Physical Activity Level,
Sleep Duration, Gender, Age, and Occupation. The other variables were
determined to have too weak of a relationship after our EDA and were
therefore left out of the analyses.

The Occupation variable was then transformed to group the various
occupations into the following groups: Healthcare, Education,
Engineering, Business/Finance, and Science. Prior to performing any
subsequent analyses, we removed the original versions of the variables
we transformed: systolic, diastolic, daily steps, occupation,
Physical.Activity.Level, Heart.Rate, and AgeGroup.

The objective of our analysis was to determine which of these variables
are most important for determining an individual's sleep quality and
what specific effect they have. After conducting an EDA we hypothesized
that stress, sleep duration, and occupation would have the largest
impact on sleep quality. We believed that stress and sleep quality would
be negatively correlated, sleep duration and quality would be positively
correlated, and that jobs with longer or more variable hours would
negatively affect sleep quality.

## Model Development

Random Forest Models and 5-fold Cross-Validation were considered to be
the "best" fit for our outcome of interest, Quality of Sleep. The data
were not split into train/test subsets as the data contains less than
400 observations and many of the values were unique in comparison to the
rest of the data. Prior to constructing this model, a colinearity plot
was constructed to determine any presence of colinearity in our data.

![Figure 1: A colinearity plot demonstrating the colinearity between our
variables.](images/collinearity.png){width="586"}

Stress and Sleep Quality had an absolute value of 0.9 on the correlation
scale. Sleep Duration and Stress were also seen to be strongly
correlated. Stress was still included in our models due to its
well-studied impact on sleep; however, we found it important to note the
strong linear relationship between Stress and Sleep Quality.

## Defining the Models

We fit GLMs for each predictor in order to determine their effect on
sleep quality. Each model only used a single predictor as the only aim
was to quantify and understand the impact of each for our outcome of
interest. For our Random Forest model, we used mtry (with a range of
1-7) and min_n (with a range of 1-21) as tuning parameters to create 300
different decision trees. Mtry controls how many predictors a decision
tree can consider at any given point in time. It adds randomness to the
decision tree creation process and ensures that all of the trees do not
look the same. Min_n specifies the minimum number of samples that should
be present in the leaf node after splitting a node. In our case, this
means that if any terminal node has more than 21 observations and is not
a pure node, we can split it further into subnodes. This was all done
using 5-fold Cross-Validation repeated 5 times in order to train the
data to predict on an unseen dataset.

## Evaluation of Models

Root Mean Squared Error (RMSE) is a common metric used to evaluate
regression models. The RMSE is formally defined as the square root of
the mean square of all error and is defined by the following formula:
$$\sum_{i=1}^{D}(x_i-y_i)^2$$

It is important to consider that while RMSE is scale dependent, common
practice notes that low RMSE values indicate stronger model performance.

R-squared (R²) is a statistical measure used in regression analyses. It
quantifies the proportion of the variance in a dependent variable that
can be explained by an independent variable within a regression model
and is calculated with the following formula:

$$
[ R^2 = 1 - \frac{{\text{{Unexplained Variation}}}}{{\text{{Total Variation}}}} ]
$$

Null models provide a baseline for the evaluation of a model's
performance. They generate predictions for an outcome of interest
without using any predictor variables. Instead, they simply average the
values of the outcome of interest to make these "predictions". We used
both RMSE and R² values to evaluate the performance of our model
compared to a "null model".

## Software Used for Analyses

This analysis was conducted under R version 4.3.1 on a MacOS operating
system. The following R packages were used in the development of these
analyses: here, skimr, broom, tidyverse, ggplot2, dplyr, corrplot,
ranger, and vip. All processing and analysis code can be found in the
Supplementary Material file.

# Results (4)

## Outcomes of Interest

The 'Quality of Sleep' variable has been selected as our outcome of
interest. It is important to note that this variable is a subjective
measure of a participant's self-reported sleep quality. Figure 2 shows
that most subjects reported a sleep quality score of 8 (out of 10).

![Figure 2: Distribution of Sleep Quality Scores in the
dataset.](images/unnamed-chunk-17-1.png)

Our Random Forest Model built with 5-fold Cross-Validation explored
different combinations of predictors to determine which parameters gave
the most accurate predictions of sleep quality.

![Figure 3: The results generated from our RF model. RMSE increases when
more predictors are added to the model. The R-squared value also
decreases as more predictors are added to the
model.](images/RFtuningresults.png)

Figure 3 shows that as we include more predictors in our model, RMSE
increases slightly while the R-squared value decreases. This indicates
that some variables in our dataset are more important for predicting
sleep quality than the others. Based off of our variable importance
graph, we can assume that stress level, sleep duration, and age are the
factors that carry the most weight when predicting sleep quality.

We went on to evaluate how well our RF model predicted values by
creating an observed vs. predicted plot as seen in figure 4.

![Figure 4: A comparison of the actual values for sleep quality in our
dataset (observed) compared to the values predicted by our fitted RF
model.](images/observed_vs_predicted.png)

While the predicted values don't align perfectly with the red "ideal
value" line, they fall very close overall, indicating that our model can
sufficiently predict sleep quality using unseen data.

![Figure 5: The observed vs. predicted values for our "Null
model".](images/null_observedvspredicted.png)

We can see in figure 5 that the null model only predicts values that are
the average of our observed values. As expected, these predictions are
far from our actual observed values.

![Table 1: The results of our GLMs. The "term" column identifies the
category/variable being analyzed. The "model" column identifies the
predictors used to fit each model. The "statistic" column is a relative
measure of the strength of the relationship between each predictor and
sleep quality. The "estimate" column shows the predicted change in sleep
quality score as a result of each associated
predictor.](images/table.png)

We believed that stress and sleep quality would be negatively
correlated, sleep duration and quality would be positively correlated,
and that jobs with longer or more variable hours would negatively affect
sleep quality. As seen in table 1, we found that all of our assumptions
were supported. The "intercept" values under the term column represents
what the expected sleep quality score would be if a continuous predictor
was at a value of 0. For categorical variables, it represents the
reference category (e.g. the intercept for the Gender model represents
the expected sleep quality score for females).

## Machine Learning Models

A variable importance plot (VIP) (figure 6) was constructed using the
"best" set of parameters (as guided by our RF model) to determine which
variables had the strongest impact on sleep quality.

![Figure 6: The ranking of each variable's importance in determining
sleep quality according to our Random Forest
model.](images/variableimportance.png)

As seen in figure 6, Sleep Duration, Stress Level, and Age had the
highest levels of importance in our model. Physical Activity Level was
excluded entirely, which was to be expected given our exploratory data
analysis and simple linear regression models.

![Table 2: A comparison of the RMSE and R-squared values generated by
the null and RF models. The null model has no R-squared value due to its
method of generating predictions.](images/metric_table.png){width="510"}

As seen in table 2, our null model was found to have an RMSE of 1.193;
compared to our best parameter RF model's RMSE of 0.059, this is
relatively high. The R-squared value doesn't exist for the null model
since there is technically no variation between the observed and
predicted values.

# Discussion (5)

## Summary

The differences between the null model and RF model's metrics show that
our RF model is better suited for predicting sleep quality on unseen
data than a model using no predictors. The R-squared value of 0.997 is
extremely high; this shows that a large amount of variation is explained
using our RF model. The variable importance plot we created answers our
questions about which variables are most important for predicting sleep
quality. As we can see, sleep duration, stress level, and age were the
top three predictors for sleep quality in our data. This reflects the
relationships we saw in our colinearity plot as well as the clear
correlations we found between the variables in our EDA. We did not
expect to see age so high on the list; this could be due to us grouping
ages together in our EDA and therefore misleading us in our preliminary
data exploration. We did not use the groups when creating our models as
we learned that grouping continuous variables is a poor practice as we
progressed in our MADA course.

Through our linear models, we determined that our hypotheses regarding
each variable's effect on quality of sleep were supported. The variables
determined to be the most important by our RF model had the following
effects on sleep quality: sleep duration was positively correlated, age
was positively correlated, and stress level was negatively correlated.

## Strengths and Limitations

The main limitations of our models stemmed from the data itself. As
stated in our introduction section, we have been unable to identify
exactly where the data were collected and how it was gathered. The
values for stress level and sleep duration were also based on a
subjective rating given by each participant in the study; this could
introduce individual bias to the data and prevent an objective
evaluation of the effect of these variables. An inherent limitation in
RF models is difficulty of interpretation; we attempted to minimize this
by creating easy to read tables and graphs for the results.
We chose not to split our data into a train/test split due to the small number of observations and presence of rare values in our data. We believe that fitting using CV minimized the downsides of this. Future
analyses would benefit from more detailed meta data and a better
understanding of the experimental design that generated our
observations.

The strengths of our analysis lie in the easily interpretation of our
GLMs and the predictive power of an RF model. We were able to clearly
display and quantify the relationships between our predictors and sleep
quality, answering our question about each variable's effect on our
outcome of interest. We were also able to create a powerful RF model
that explained nearly all of the variation in our data as well as a
ranking of each variable's impact on sleep quality. Fitting our RF model
with CV allowed us to train the model to create predictions for "unseen"
data as well.

## Conclusion

Overall, while our study limitations prevent us from generalizing the
findings of this analysis broadly, we were able to clearly answer our
questions about the data. Our findings supported our hypotheses about
the effects of each predictor on sleep quality. However, our hypothesis
about which variables would be the most important was not supported by
the analysis. Instead, we determined that stress level, sleep duration,
and age were the most important predictors for our outcome of interest.
Ultimately, the results indicate that the sleep quality of this study's
participants is impacted disproportionately by an assortment of
variables dealing with physical, mental, and environmental factors.

{{< pagebreak >}}

# References